---
title: "olympic_dataset_report"
output: html_document
date: "2022-10-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Attaching the required libraries
```{r echo=FALSE}
library(tidyverse)
library(skimr)
library(janitor)
library(tidyr)
library(DataExplorer)
library(ggplot2)
library(reactable)
library(GGally)
library(dplyr)
```

# Step1 (Data exploring and EDA):
## We will start by importing the dataset
```{r}
olympic <- read.csv("data/athlete_events.csv")
```

# We started by exploring our dataset
```{r}
#take a look of our data
glimpse(olympic)
skim(olympic)
summary(olympic)
str(olympic)
```

# Then generated an EDA report using DataExplorer
```{r eval = FALSE}
create_report(olympic, 
              output_file = "olympics_EDA_report.html", 
              output_dir = "EDA_reports/")
```


# Step2 (Data cleaning):

### We figured out that we need to do column name cleaning using janitor package
```{r}
olympic_df <- clean_names(olympic)
```

## Then, modified the data type to be factor
```{r}
olympic_df$sex <- as.factor(olympic_df$sex)
olympic_df$season <- as.factor(olympic_df$season)
olympic_df$city <- as.factor(olympic_df$city)
olympic_df$sport <- as.factor(olympic_df$sport)
olympic_df$event <- as.factor(olympic_df$event)
olympic_df$country <- as.factor(olympic_df$team)
olympic_df$noc <- as.factor(olympic_df$noc)
```

## We decided to change some column names to make it more clear
```{r}
#rename team column with country
olympic_df%>%
  select(team) %>%
  rename("country"= "team")
```



## Our dataset has missing values in the medal column. Thus, we decided to deal with them
```{r}
# print number of missing values
#print(sum(is.na(olympic_df)))
```

```{r}
#sum(is.na(olympic_df)) 
#remove NULL value in medal 
#the NULL mean not winning any medals
#olympic_df %>% 
#  group_by(medal) %>% 
#  count(medal)

#olympic_df$medal <- olympic_df$medal[is.na(olympic_df$medal)]="Not win"
#table(olympic_df$medal)

#modify type of data
#olympic_df$medal <- as.factor(olympic_df$medal)
```

