---
title: "presentation"
output:
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
date: "2022-10-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Attaching the required libraries 

```{r echo=FALSE}
library(tidyverse)
library(skimr)
library(janitor)
library(tidyr)
library(DataExplorer)
library(ggplot2)
library(reactable)
library(GGally)
library(dplyr)
```

# Step1 (Data exploring and EDA):

## We will start by importing the dataset

```{r}
olympic <- read.csv("data/athlete_events.csv")
```

# We started by exploring our dataset

```{r}
#take a look of our data
glimpse(olympic)
skim(olympic)
summary(olympic)
str(olympic)
```

# Then generated an EDA report using DataExplorer

```{r eval = FALSE}
create_report(olympic, 
              output_file = "olympics_EDA_report.html", 
              output_dir = "EDA_reports/")
```

# Step2 (Data cleaning):

### We figured out that we need to do column name cleaning using janitor package

```{r}
olympic_df <- clean_names(olympic)
```

## Then, modified the data type to be factor

```{r}
olympic_df$sex <- as.factor(olympic_df$sex)
olympic_df$season <- as.factor(olympic_df$season)
olympic_df$city <- as.factor(olympic_df$city)
olympic_df$sport <- as.factor(olympic_df$sport)
olympic_df$event <- as.factor(olympic_df$event)
olympic_df$country <- as.factor(olympic_df$team)
olympic_df$noc <- as.factor(olympic_df$noc)
```

## Our dataset has missing values in the medal column. Thus, we decided to deal with them

```{r}
sum(is.na(olympic_df)) 

# remove NULL value in medal 
# the NULL mean not winning any medals
olympic_df %>% 
  group_by(olympic_df$medal) %>% 
  count(olympic_df$medal)

olympic_df$medal[is.na(olympic_df$medal)] = "Not win"
table(olympic_df$medal)

#modify type of data
olympic_df$medal <- as.factor(olympic_df$medal)
```

# Step 3 (find answers to our questions):

### Q1: What is the total number of sporting events (for winter and summer)?

```{r}
#Total number of sporting events (for winter and summer)

#Unique Olympic sports
unique(olympic_df$sport)

#Number of Unique Sports
length(unique(olympic_df$sport))
```

### Q2: What is the total number of sporting events (for winter and summer)?

```{r}
#Table that shows the total number of athletes and events each year
count_year_season <- olympic_df %>%
  group_by(year, season) %>%
  summarize(
    Athletes = length(unique(id)),
    Events = length(unique(event))
  )
count_year_season 


#Line plot that shows the number of sporting events over time
count_year_season %>% 
  ggplot(count_year_season, mapping = aes(x=year, y=Events,color=season)) + 
  ggtitle("Number of sporting events over time") +
  geom_line() 

#Line plot that shows the number of athleets events over time
count_year_season %>% 
  ggplot(count_year_season, mapping = aes(x=year, y=Athletes,color=season)) + 
  ggtitle("Number of athleets over time") +
  geom_line() 
```

### Q3: What is the total number of medals won by each country?

```{r}
country_medal_count <- olympic_df %>%
  filter(medal !='Not win')%>%
  group_by(noc,medal) %>% 
  summarise(Total = n())

country_medal_count 

country_medal_count %>%
  ggplot(country_medal_count,mapping = aes(x=noc, y=Total, fill=medal)) +
  scale_fill_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) +
  geom_col() +
  ggtitle("Total number of Medal won by each country") 

by_country <- olympic_df %>%filter(medal !='Not win')%>% group_by(team) %>% summarise(Total = n())
by_country
```

### Q4: Who are the top 3 countries that have the most number of participants over the years and how do we compare them over the years?

```{r}
head(sort(table(olympic_df$noc), TRUE),5)
```

#### United States:

```{r}
# USA has 18853 participants
# USA won 1358 bronze; 2638 gold; 1641 silver = 5637
USA_won_medals <- data.frame(olympic_df[olympic_df$noc == 'USA' & olympic_df$medal != 'Not win',])

USA_won_medals %>%
  select(medal, year) %>%
  group_by(medal) %>%
  count(year) -> USA_won_medals

#table(USA_won_medals$year)

ggplot(data = USA_won_medals,
       mapping = aes(x=year, y=n, color=medal)) + 
  scale_color_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) +
  ggtitle("Medals won over the years for USA") +
  geom_line() + facet_grid(rows = vars(medal))
```

#### France

```{r}
# France has 12758 participants
# France won 666 bronze; 501 gold; 610 silver = 1777
FRA_won_medals <- data.frame(olympic_df[olympic_df$noc == 'FRA' & olympic_df$medal != 'Not win',])

#table(FRA_won_medals$year)

FRA_won_medals %>%
  select(medal, year) %>%
  group_by(medal) %>%
  count(year) -> FRA_won_medals

ggplot(data = FRA_won_medals,
       mapping = aes(x=year, y=n, color=medal)) + 
  scale_color_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) + 
  ggtitle("Medals won over the years for France") +
  geom_line() + facet_grid(rows = vars(medal)) 
```

#### United Kingdom:

```{r}
# Great Brittin has 12256 participants
# Great Brittin won 651 bronze; 678 gold; 739 silver = 2068
GBR_won_medals <- data.frame(olympic_df[olympic_df$noc == 'GBR' & olympic_df$medal != 'Not win',])

#table(GBR_won_medals$year)

GBR_won_medals %>%
  select(medal, year) %>%
  group_by(medal) %>%
  count(year) -> GBR_won_medals

ggplot(data = GBR_won_medals,
       mapping = aes(x=year, y=n, color=medal)) + 
  scale_color_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) + 
  ggtitle("Medals won over the years for Great Brittin") +
  geom_line() + facet_grid(rows = vars(medal)) 
```

### Q5: What is the percentage of the distribution of male and female athletes throughout the years of Olympic history?

```{r}
#Percentage of the distribution of male and female athletes throughout the years of Olympic history

Gender_Dist <- olympic_df %>%
  group_by(year, sex) %>%
  summarise(total_count=n()) %>% 
  mutate(total_Frequency_count = (total_count / sum(total_count) )*100)

ggplot(Gender_Dist, aes(x=year, y=total_Frequency_count ,color=sex)) + 
  geom_point() 
```

### Q6: How many male and females participated in the Olympics?

```{r}
#How many male and females?
olympic_df%>%
  group_by(sex)%>%
  summarise(count= length(unique(id))) -> gender_count

gender_count

gender_count %>%
  ggplot(aes(x=sex,y= count,fill=sex)) +
  ggtitle("Number of Male and Female participants over the years") +
  geom_col()
```

### Q7: Can we comparing men and women winners over the years for summer and winter?

#### Summer:

```{r}
olympic_df %>% 
  filter(medal != "Not win", season == "Summer") %>% 
  group_by(year , sex , medal, season) %>%
  summarise(contestants = length(unique(id))) -> FVSM

ggplot(FVSM ,aes(x = year, y = contestants ,color = sex))+
  geom_point()+
  geom_line( )+
  ggtitle("Summer Men and Womens winners over the years") +
  theme(plot.title = element_text(hjust = 0.5))+
  facet_wrap(FVSM$medal)
```

#### Winter:

```{r}
olympic_df %>% 
  filter(medal != "Not win", season == "Winter") %>% 
  group_by(year , sex , medal, season) %>%
  summarise(contestants = length(unique(id))) -> FVSMWinter

ggplot(FVSMWinter ,aes(x = year, y = contestants ,color = sex))+
  geom_point()+
  geom_line( )+
  ggtitle("Winter Men and Womens winners over the years") +
  theme(plot.title = element_text(hjust = 0.5))+
  facet_wrap(FVSMWinter$medal)
```

### Q8: How many medal have been won by male and female?

```{r}
#How many medal have been won by male and female?
medal_counts <-olympic_df%>%
  filter(medal!= "Not win") %>% 
  group_by(sex) %>% 
  summarize(Count= length(medal), .groups='drop')

olympic_df%>%
  group_by(sex, medal)%>%
  count(medal)

medal_counts <- olympic_df %>% 
  filter(medal!= "Not win") %>%
  group_by(sex,medal) %>% 
  summarize(Count = length(medal), .groups = 'drop') %>% 
  ggplot(aes(x = sex, y = Count , fill = medal)) + 
  scale_fill_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) +
  ggtitle("Number of males and females who won medals")+
  geom_col(position="dodge") 

medal_counts
```

### Q9: What is the most productive age for both male and female athletes in all sorts of sports?

```{r}
olympic_df %>%
  filter(age >= 10 , age <= 75) %>% 
  group_by(age, sex) %>% 
  summarise(contestants = length(unique(id))) -> fvsminage


ggplot(fvsminage ,aes(x =age ,y = contestants))+
  geom_point()+
  #stat_summary( fun  = "mean")+
  geom_vline(aes(xintercept =23))+
  ggtitle("The most productive age for both male and female athletes") +
  theme(plot.title = element_text(hjust = 0.5))+
  facet_wrap(fvsminage$sex)
```

### Q10: Is there a relation between athletes' gender and their winning to their olympic game?

```{r}
olympic_df %>% 
  filter(medal != "Not win") %>%
  group_by(sex , age) %>%
  summarise(contestants = length(unique(id))) -> fvsminevent


fvsminevent %>% 
  arrange(-contestants , n =15) %>% 
  head(15L) -> topevents


olympic_df %>% 
  filter(medal != "Not win", sex == 'F') %>%
  group_by(sex , event , medal) %>%
  summarise(contestants = length(unique(id))) -> Finevent

Finevent %>% 
  arrange(-contestants , n =10) %>% 
  head(30L) -> topFinevent

olympic_df %>% 
  filter(medal != "Not win", sex == 'M') %>%
  group_by(sex , event , medal) %>%
  summarise(contestants = length(unique(id))) -> Minevent

Minevent %>% 
  arrange(-contestants , n =10) %>% 
  head(30L) -> topMinevent


topFinevent %>% 
  ggplot(aes(x =contestants,y=fct_reorder(event,contestants) , fill = medal))+
  geom_col()+
  ggtitle("The most winning events for womens") +
  ylab('Event name')+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = c("#8B3E2F",'#CD9B1D','#C1CDC1'))


topMinevent %>% 
  ggplot(aes(x =contestants,y=fct_reorder(event,contestants) , fill = medal))+
  geom_col()+
  ggtitle("The most winning events for men") +
  ylab('Event name')+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = c("#8B3E2F",'#CD9B1D','#C1CDC1'))
```

# Step 4 (Hypothesis testing questions):

### Women growth since they joined the olympics

```{r}
#women grouth
olympic_df %>% 
  filter(sex =='F') %>% 
  group_by(year , sex ,season) %>%
  summarise(contestants = length(unique(id))) -> Fgrouth


ggplot(Fgrouth ,aes(x = year, y = contestants))+
  geom_point()+
  geom_line()+
  ggtitle("Womens contestants through years") +
  theme(plot.title = element_text(hjust = 0.5))+
  facet_grid(Fgrouth$season)
```

### Men growth vs. women growth

```{r}
#MEN GROUTH VS WOMENS
olympic_df %>%
 # filter(season == 'Summer') %>% 
  group_by(year , sex, season ) %>%
  summarise(contestants = length(unique(id))) -> FVSMinyears


ggplot(FVSMinyears ,aes(x = year, y = contestants ,color = sex ))+
  geom_point()+
  geom_line()+
  ggtitle("Mens and Womens contestants through years in both seasons") +
  theme(plot.title = element_text(hjust = 0.5))+
  facet_grid(FVSMinyears$season)
```

# Step 5 (Summary):
