---
title: "Olympic Dataset Presentation"
format: 
  revealjs:
    slide-number: true
    theme: [solarized]
editor: visual
---

## Presented by: {auto-animate="true"}

::: {.fragment .fade-up}
-   Mukhtar Alhejji
:::

::: {.fragment .fade-up}
-   Noura Bin Salamah
:::

::: {.fragment .fade-up}
-   Abdulrahman Almegren
:::

::: {.fragment .fade-up}
-   Razan Sendi
:::

## Attach required libraries

```{r echo=TRUE}
library(tidyverse)
library(skimr)
library(janitor)
library(tidyr)
library(DataExplorer)
library(ggplot2)
library(reactable)
library(GGally)
library(dplyr)
```

# Step1 (Data exploring and EDA)

## Reading the dataset

```{r echo=TRUE}
olympic <- read.csv("data/athlete_events.csv")
```

## Exploring the dataset

::: panel-tabset
### Head

```{r}
head(olympic, 8)
```

### Glimpse

```{r}
glimpse(olympic)
```

### Summary

```{r}
summary(olympic)
```

### Str

```{r}
str(olympic)
```
:::

```{r}
#take a look of our data
#glimpse(olympic)
#skim(olympic)
#summary(olympic)
#str(olympic)
```

## Generating EDA report with DataExplorer

```{r eval = FALSE}
create_report(olympic, 
              output_file = "olympics_EDA_report.html", 
              output_dir = "EDA_reports/")
```

# Step2 (Data cleaning)

## Column names cleaning with janitor

```{r echo=TRUE}
olympic_df <- clean_names(olympic)
```

### Modified data type to be factor

```{r echo=TRUE}
olympic_df$sex <- as.factor(olympic_df$sex)
olympic_df$season <- as.factor(olympic_df$season)
olympic_df$city <- as.factor(olympic_df$city)
olympic_df$sport <- as.factor(olympic_df$sport)
olympic_df$event <- as.factor(olympic_df$event)
olympic_df$country <- as.factor(olympic_df$team)
olympic_df$noc <- as.factor(olympic_df$noc)
```

## Missing values

```{r}
sum(is.na(olympic_df)) 

# remove NULL value in medal 
# the NULL mean not winning any medals
olympic_df %>% 
  group_by(olympic_df$medal) %>% 
  count(olympic_df$medal)

olympic_df$medal[is.na(olympic_df$medal)] = "Not win"
table(olympic_df$medal)

#modify type of data
olympic_df$medal <- as.factor(olympic_df$medal)
```

# Step 3 (find answers to our questions)

------------------------------------------------------------------------

## Q1: What is the total number of sporting events (for winter and summer)?

```{r echo=TRUE}
#Number of Unique Sports
length(unique(olympic_df$sport))

#Unique Olympic sports
head(unique(olympic_df$sport), 25)
```

## Q2: What is the total number of athletes and sporting events (for winter and summer)?

```{r echo=FALSE}
#Table that shows the total number of athletes and events each year
count_year_season <- olympic_df %>%
  group_by(year, season) %>%
  summarize(
    Athletes = length(unique(id)),
    Events = length(unique(event))
  )
#count_year_season 
```

```{r}
#Line plot that shows the number of sporting events over time
count_year_season %>% 
  ggplot(count_year_season, mapping = aes(x=year, y=Events,color=season)) + 
  ggtitle("Number of events over time") +
  geom_line() 
```

## Q2: What is the total number of athletes and sporting events (for winter and summer)?

```{r}
#Line plot that shows the number of athleets events over time
count_year_season %>% 
  ggplot(count_year_season, mapping = aes(x=year, y=Athletes,color=season)) + 
  ggtitle("Number of athleets over time") +
  geom_line() 
```

## Q3: What is the top countries who won meadls?

```{r}
country_medal_count <- olympic_df %>%
  filter(medal !='Not win')%>%
  group_by(noc,medal) %>% 
  summarise(Total = n())

#country_medal_count 
```

```{r}
country_medal_count %>% 
  arrange(-Total , n =50) %>% 
  head(30L) -> top_medal

top_medal %>%
  ggplot(mapping = aes(x=fct_reorder(noc,Total), y=Total, fill=medal)) +
  xlab("Country") +
  scale_fill_manual(values =c("#8B3E2F",'#CD9B1D','#8B8386'))+
  geom_col() +
  ggtitle("Top countries won medals")+
  theme(axis.text.x = element_text(angle = 45)) 

#country_medal_count %>%
#  ggplot(country_medal_count,mapping = aes(x=noc, y=Total, fill=medal)) +
#  scale_fill_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) +
#  geom_col() +
#  ggtitle("Total number of Medal won by each country") 
```

```{r echo=FALSE}
by_country <- olympic_df %>%filter(medal !='Not win')%>% group_by(team) %>% summarise(Total = n())
#by_country
```

## Q4: What are the top 3 countries that have the most number of participants over the years?

```{r echo=TRUE}
head(sort(table(olympic_df$noc), TRUE),5)
```

## Top 3 countries performance:

::: panel-tabset
### United States in Summer

```{r}
# USA has 18853 participants
# USA won 1358 bronze; 2638 gold; 1641 silver = 5637
USA_won_medals <- data.frame(olympic_df[olympic_df$noc == 'USA' & olympic_df$medal != 'Not win'  & olympic_df$season == 'Summer',])

USA_won_medals %>%
  select(medal, year) %>%
  group_by(medal) %>%
  count(year) -> USA_won_medals

#table(USA_won_medals$year)

ggplot(data = USA_won_medals,
       mapping = aes(x=year, y=n, color=medal)) + 
  scale_color_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) +
  ggtitle("Summer Medals won over the years for USA") +
  geom_line() + facet_grid(rows = vars(medal))
```

### United States in Winter

```{r}
# USA has 18853 participants
# USA won 1358 bronze; 2638 gold; 1641 silver = 5637
USA_won_medals <- data.frame(olympic_df[olympic_df$noc == 'USA' & olympic_df$medal != 'Not win'  & olympic_df$season == 'Winter',])

USA_won_medals %>%
  select(medal, year) %>%
  group_by(medal) %>%
  count(year) -> USA_won_medals

#table(USA_won_medals$year)

ggplot(data = USA_won_medals,
       mapping = aes(x=year, y=n, color=medal)) + 
  scale_color_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) +
  ggtitle("Winter Medals won over the years for USA") +
  geom_line() + facet_grid(rows = vars(medal))
```

### France in Summer

```{r}
# France has 12758 participants
# France won 666 bronze; 501 gold; 610 silver = 1777
FRA_won_medals <- data.frame(olympic_df[olympic_df$noc == 'FRA' & olympic_df$medal != 'Not win' & olympic_df$season == 'Summer',])

#table(FRA_won_medals$year)

FRA_won_medals %>%
  select(medal, year) %>%
  group_by(medal) %>%
  count(year) -> FRA_won_medals

ggplot(data = FRA_won_medals,
       mapping = aes(x=year, y=n, color=medal)) + 
  scale_color_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) + 
  ggtitle("Summer Medals won over the years for France") +
  geom_line() + facet_grid(rows = vars(medal)) 
```

### France in Winter

```{r}
# France has 12758 participants
# France won 666 bronze; 501 gold; 610 silver = 1777
FRA_won_medals <- data.frame(olympic_df[olympic_df$noc == 'FRA' & olympic_df$medal != 'Not win' & olympic_df$season == 'Winter',])

#table(FRA_won_medals$year)

FRA_won_medals %>%
  select(medal, year) %>%
  group_by(medal) %>%
  count(year) -> FRA_won_medals

ggplot(data = FRA_won_medals,
       mapping = aes(x=year, y=n, color=medal)) + 
  scale_color_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) + 
  ggtitle("Winter Medals won over the years for France") +
  geom_line() + facet_grid(rows = vars(medal)) 
```

### United Kingdom in Summer

```{r}
# Great Brittin has 12256 participants
# Great Brittin won 651 bronze; 678 gold; 739 silver = 2068
GBR_won_medals <- data.frame(olympic_df[olympic_df$noc == 'GBR' & olympic_df$medal != 'Not win' & olympic_df$season == 'Summer',])

#table(GBR_won_medals$year)

GBR_won_medals %>%
  select(medal, year) %>%
  group_by(medal) %>%
  count(year) -> GBR_won_medals

ggplot(data = GBR_won_medals,
       mapping = aes(x=year, y=n, color=medal)) + 
  scale_color_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) + 
  ggtitle("Summer Medals won over the years for Great Brittin") +
  geom_line() + facet_grid(rows = vars(medal)) 
```

### United Kingdom in Winter

```{r}
# Great Brittin has 12256 participants
# Great Brittin won 651 bronze; 678 gold; 739 silver = 2068
GBR_won_medals <- data.frame(olympic_df[olympic_df$noc == 'GBR' & olympic_df$medal != 'Not win' & olympic_df$season == 'Winter',])

#table(GBR_won_medals$year)

GBR_won_medals %>%
  select(medal, year) %>%
  group_by(medal) %>%
  count(year) -> GBR_won_medals

ggplot(data = GBR_won_medals,
       mapping = aes(x=year, y=n, color=medal)) + 
  scale_color_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) + 
  ggtitle("Winter Medals won over the years for Great Brittin") +
  geom_line() + facet_grid(rows = vars(medal)) 
```
:::

## Q5: How many male and females participated in the Olympics?

```{r}
#How many male and females?
olympic_df%>%
  group_by(sex)%>%
  summarise(count= length(unique(id))) -> gender_count

#gender_count

gender_count %>%
  ggplot(aes(x=sex,y= count,fill=sex)) +
  ggtitle("Number of Male and Female participants") +
  geom_col()
```

## Q6: What is the percentage of male vs. female athletes throughout the years of Olympic history?

```{r}
#Percentage of the distribution of male and female athletes throughout the years of Olympic history

Gender_Dist <- olympic_df %>%
  group_by(year, sex) %>%
  summarise(total_count=n()) %>% 
  mutate(total_Frequency_count = (total_count / sum(total_count) )*100)

ggplot(Gender_Dist, aes(x=year, y=total_Frequency_count ,color=sex)) + 
  geom_point() 
```

## Q7: Can we compare men and women winners over the years for summer and winter?

## Q7: Can we compare men and women winners ...

::: panel-tabset
### Summer

```{r}
olympic_df %>% 
  filter(medal != "Not win", season == "Summer") %>% 
  group_by(year , sex , medal, season) %>%
  summarise(contestants = length(unique(id))) -> FVSM

ggplot(FVSM ,aes(x = year, y = contestants ,color = sex))+
  geom_point()+
  geom_line( )+
  ggtitle("Summer Men and Womens winners over the years") +
  theme(plot.title = element_text(hjust = 0.5))+
  facet_wrap(FVSM$medal)
```

### Winter:

```{r}
olympic_df %>% 
  filter(medal != "Not win", season == "Winter") %>% 
  group_by(year , sex , medal, season) %>%
  summarise(contestants = length(unique(id))) -> FVSMWinter

ggplot(FVSMWinter ,aes(x = year, y = contestants ,color = sex))+
  geom_point()+
  geom_line( )+
  ggtitle("Winter Men and Womens winners over the years") +
  theme(plot.title = element_text(hjust = 0.5))+
  facet_wrap(FVSMWinter$medal)
```
:::

## Q8: How many medal have been won by male vs. female?

```{r}
#How many medal have been won by male and female?
medal_counts <-olympic_df%>%
  filter(medal!= "Not win") %>% 
  group_by(sex) %>% 
  summarize(Count= length(medal), .groups='drop')

#olympic_df%>%
#  group_by(sex, medal)%>%
#  count(medal)

medal_counts <- olympic_df %>% 
  filter(medal!= "Not win") %>%
  group_by(sex,medal) %>% 
  summarize(Count = length(medal), .groups = 'drop') %>% 
  ggplot(aes(x = sex, y = Count , fill = medal)) + 
  scale_fill_manual(values = c("#8B3E2F",'#CD9B1D','#8B8386')) +
  ggtitle("Number of males and females who won medals")+
  geom_col(position="dodge") 

medal_counts
```

## Q9: What is the average age for both male and female athletes in all sorts of sports?

```{r}
olympic_df %>%
  filter(age >= 10 , age <= 75) %>% 
  group_by(age, sex) %>% 
  summarise(contestants = length(unique(id))) -> fvsminage


ggplot(fvsminage ,aes(x =age ,y = contestants))+
  geom_point()+
  #stat_summary( fun  = "mean")+
  geom_vline(aes(xintercept =23))+
  ggtitle("The average age for both male and female athletes") +
  theme(plot.title = element_text(hjust = 0.5))+
  facet_wrap(fvsminage$sex)
```

## Q10: Is there a relation between athletes' gender and their winning to their olympic game?

```{r}
olympic_df %>% 
  filter(medal != "Not win") %>%
  group_by(sex , age) %>%
  summarise(total = n() ) -> fvsminevent


fvsminevent %>% 
  arrange(-total , n =15) %>% 
  head(15L) -> topevents


olympic_df %>% 
  filter(medal != "Not win", sex == 'F') %>%
  group_by(sex , sport , medal) %>%
  summarise(total = n() ) -> Finevent

Finevent %>% 
  arrange(-total , n =10) %>% 
  head(30L) -> topFinevent

olympic_df %>% 
  filter(medal != "Not win", sex == 'M') %>%
  group_by(sex , sport , medal) %>%
  summarise(total = n() ) -> Minevent

Minevent %>% 
  arrange(-total , n =10) %>% 
  head(48L) -> topMinevent


```

## Q10: Is there a relation between athletes' ...

::: panel-tabset
### Women

```{r}
topFinevent %>% 
  ggplot(aes(x =total,y=fct_reorder(sport,total) , fill = medal))+
  geom_col()+
  ggtitle("The most winning events for womens") +
  ylab('Event name')+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = c("#8B3E2F",'#CD9B1D','#C1CDC1'))

```

### Men

```{r}
topMinevent %>% 
  ggplot(aes(x =total,y=fct_reorder(sport,total) , fill = medal))+
  geom_col()+
  ggtitle("The most winning events for men") +
  ylab('Event name')+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = c("#8B3E2F",'#CD9B1D','#C1CDC1'))
```
:::

# Step 4 (Hypothesis testing questions)

## Women growth since they joined the olympics

```{r}
#women grouth
olympic_df %>% 
  filter(sex =='F') %>% 
  group_by(year , sex ,season) %>%
  summarise(contestants = length(unique(id))) -> Fgrouth


ggplot(Fgrouth ,aes(x = year, y = contestants))+
  geom_point()+
  geom_line()+
  ggtitle("Womens contestants through years") +
  theme(plot.title = element_text(hjust = 0.5))+
  facet_grid(Fgrouth$season)
```

## Men and women constantly growth

```{r}
#MEN GROUTH VS WOMENS
olympic_df %>%
 # filter(season == 'Summer') %>% 
  group_by(year , sex, season ) %>%
  summarise(contestants = length(unique(id))) -> FVSMinyears


ggplot(FVSMinyears ,aes(x = year, y = contestants ,color = sex ))+
  geom_point()+
  geom_line()+
  ggtitle("Mens and Womens contestants through years in both seasons") +
  theme(plot.title = element_text(hjust = 0.5))+
  facet_grid(FVSMinyears$season)
```

# Step 5 (Summary)

## Summary

::: {.fragment .fade-left}
-   While observing the data, a few major drops in the number of athletes and number of countries participating in the Olympic games was overseen which posed as a mystery where such decrease in the number of participants was due to a boycott that took place.
-   For the 1980s, the number of female participants during the Olympic Summer Games has gradually increased and the total number of female participation during such multi-sport event is in fact expected to keep increasing for the upcoming years.
:::

## Summary ...

-   Irregularities in the number of male participants was overseen in our analysis where is come years the participation of males increased and other years where it decreases (correlated to the Olympic Boycotts)
-   An external dataset that contains data about the 2021 Olympics that were taken place in Tokyo was retrieved ,where the 45 sporting events contained within the dataset was compared in regard to the male and female distribution. Hence, for mostly all the sporting events there was in fact a 50-50%precent distribution for male and female athletes.

## Summary ...

-   After observing the data , the frequency distribution of male and female athletes throughout the years of Olympic history is in fact leaning towards gender equality where the number of females participating in the Olympic games has increased during both the summer and winter games

## Questions

::: {.fragment .fade-left}
Thank you for listening, do you have any questions?
:::
